% acknowledgements.tex:

\begin{acknowledgements}

Most of all, I would like to thank my dissertation advisor, Eduardo Sontag, a talented mathematician and passionate scientist. I am proud to say my experience with him was intellectually exciting and fun, and has energized me to continue in research. I sincerely hope I continue to have opportunities to interact with Eduardo for the rest of my research career.

I am indebted to Mario Sznaier who first showed me how magic happens with control theory. I would also like to thank the members of my dissertation committee - not only for their time and extreme patience, but for their intellectual contributions to my professional development. To Carey Rappaport, who was my master thesis advisor and first taught me how to use simple mathematical models to solve complicated problems. Without the appreciation and excitement in mathematical modeling inspired by him, I may not have ever pursued research in applied mathematics. To Prof. Bahram Shafaee, I thank for being a supportive, strong guiding force as Chair of my committee. To Irina Kareva, who helped train me to work on mathemaical problems relevant to pharmaceutical industry. My experience working with Irina on chemotherapy and immunotherapy applications was extremely positive. To Mark Niedre, for having me at his lab in Fall 2015. Mark was the first person that I worked with in graduate school and I am grategul for his professional guidance.

To my groupmates, thanks for the fun and support. My experience in research was greatly enhanced as it filled out from M Ali, Tianchi, Phong, Shu, Zheming, Alon, and William. I look forward to having all of you as colleagues in the years ahead. 
\end{acknowledgements}

