\chapter*{List of Acronyms}
\addcontentsline{toc}{chapter}{List of Acronyms}

% below is the list of acronym definitions, place them in alphabetical order
% since they will not be sorted again. 
\begin{acronym}
	\acro{AUC}{Area Under the Curve}.
	The definite integral of a curve that describes the variation of a drug concentration in blood plasma as a function of time.
	
	\acro{BiTE}{Bispecific T Cell Engagers}.
	A type of fusion protein that is designed to harness the power of the immune system to treat cancer. These bispecific molecules are created by linking the targeting regions of two antibodies.
	
	\acro{IV}{Intravenous}
	Existing or taking place within, or administered into, a vein or veins
	
	\acro{KIH}{Knob Into Holes}
	A well-validated heterodimerization technology for the third constant domain of an antibody.
	
	\acro{LIGO}{Laser Interferometer Gravitational-Wave Observatory}.
	The definite integral of a curve that describes the variation of a drug concentration in blood plasma as a function of time.
	
	\acro{LTI}{Linear Time-Inveriant}
	Is a system property that produces an output signal from any input signal subject to the constraints of linearity and time-invariance.
	
	\acro{FDA}{Food and Drug Administration}.
	The United States Food and Drug Administration is a federal agency of the Department of Health and Human Services.
	
	\acro{MDOR}{Mathematically Derived Optimal Regimen}.
	
	\acro{MFI}{Maximum Fluorescence Intensity}.
	Indicates the maximum light (photons) emitted. Fluorescence is created by the absorption of energy (light) by fluorescent molecules, called fluorophores.
	
	\acro{MTD}{Maximum Tolerated Dose}.
	The maximum tolerated dose is commonly estimated to be the maximum dose that can be administered for the duration of a specific study that will not compromise the survival of the animals by causes other than carcinogenicity.

	\acro{NPI}{Nonpharmaceutical Intervention}.
	Actions, apart from getting vaccinated and taking medicine, that people and communities can take to help slow the spread of illnesses like pandemic influenza (flu).
	
	\acro{ODE}{Ordinary Differential Equation}.
	In mathematics, an ordinary differential equation is a differential equation containing one or more functions of one independent variable and the derivatives of those functions.
	
	\acro{PMP}{Pontryagin Maximum Principle}.
	Is used in optimal control theory to find the best possible control for taking a dynamical system from one state to another, especially in the presence of constraints for the state or input controls.
	
	\acro{QSS}{Quasi Steady State}.
	A situation that is changing slowly enough that it can be considered to be constant.
	
	\acro{RFM}{Ribosome Flow Model}.
	A predictive model for the fundamental features of the translation process, including translation rates, protein abundance levels, ribosomal densities and the relation between all these variables.
	
	\acro{SCID}{Severe combined immunodeficiency}
	SCID mice have a genetic immune deficiency that affects their B and T cells. Due to the lack of mature B and T lymphocytes, these mouse models are ideal for xenoengraftment of human cells and tissue.

	\acro{SISO}{Single-Input Single-Output}.
	In control theory, a single-input and single-output system is a simple single variable control system with one input and one output.
	
	\acro{SD}{Social Distancing}.
	In public health, social distancing, also called physical distancing, is a set of non-pharmaceutical interventions or measures intended to prevent the spread of a contagious disease by maintaining a physical distance between people and reducing the number of times people come into close contact with each other.
	
	\acro{TASEP}{Totally asymmetric Simple Exclusion Process}
	Is a paradigm model in nonequilibrium statistical mechanics. It is a lattice model in which particles hop into a one-dimensional lattice from the left with rate $\alpha$, jump along the lattice to the right with rate $\beta$ and exit from the right with rate 
	
	\acro{TDB}{T-cell-Dependent Bispecific}
	Similar to BiTE. A promising cancer immunotherapy that recruit a patient's T cells to kill cancer cells. 
	
	\acro{TME}{Tumor Micro-Environment}
	The ecosystem that surrounds a tumor inside the body.
	
	
\end{acronym}
